---
title: ""
output: 
  html_document:
    theme: cerulean
---
```{r setup, echo = FALSE, message = FALSE}
knitr::opts_knit$set(root.dir = '~/Documents/GitHub/tableaulpi')

library(ggplot2)
library(dplyr)
# set ggplot theme for all plots
theme_set(theme_linedraw() +
            theme(panel.grid = element_blank(),
                  strip.text = element_text(size = 14, face = "bold"),
                  axis.title = element_text(size = 12)
                  )
)
```


### What is the Living Planet Index?

Developed by Loh et al. (2005), the World Wildlife Fund (WWF), and the Zoological Society of London, the Living Planet Index (LPI) measures the overall global trend in vertebrate abundances since 1970, including terrestrial, freshwater, and marine systems. 

### How should I interpret the Living Planet Index?

```{r lpi-example-trends, echo = FALSE, fig.height = 4}
# make some simple lines
timepoints <- 1:20
df <- data.frame(
  "time" = 1989+timepoints,
  "trend" = factor(rep(c("Déclin", "Stable", "Croissance"), each = 20),
                levels = c("Déclin", "Stable", "Croissance")),
  "LPI" = c(-0.1*timepoints + 1, rep(1, 20), 0.1*timepoints + 1),
  "CI" = seq(from = 0, to = 0.5, length.out = 20)
)
# make plot
ggplot(df, aes(x = time)) +
  geom_ribbon(aes(ymin = LPI - CI, 
                  ymax = LPI + CI,
                  fill = trend), 
              alpha = 1) +
  geom_line(aes(y = LPI)) +
  geom_hline(yintercept = 1, lty = 2, lwd = .4) +
  facet_wrap(~trend) +
  labs(x = "") +
  scale_fill_manual(values = c("#f46d43", "#fee08b", "#66bd63")) +
  scale_x_continuous(breaks = c(df$time[c(1,6,11,16)])) +
  guides(fill = FALSE)
```

[ interpretation guide for each panel ]

### Limitations of the Living Planet Index

#### Represents only a small part of biodiversity

[ explanation ]

```{r richness-representativeness, echo = FALSE, warning=FALSE, fig.align="center"}
# tally number of species per taxa group
lpd_sprich <- readRDS("data/lpd_qc_fake.RDS") %>% sf::st_drop_geometry() %>%
  group_by(taxa) %>%
 dplyr::distinct(scientific_name) %>%
  tally()
lpd_sprich$category = "Suivi dans LPI"

# create fake number of total species in Quebec 
# (following Wild Species 2010 report when possible)
# https://www.canada.ca/en/environment-climate-change/services/species-risk-public-registry/publications/wild-species-2010.html
lpd_sprich <- rbind(lpd_sprich, data.frame(
  taxa = lpd_sprich$taxa,
  n = c(20, 50, 450, 100, 19),
  category = rep("Documenté au Québec", 5)
))

# set category level order
lpd_sprich$category <- factor(lpd_sprich$category, levels = rev(c("Suivi dans LPI", "Documenté au Québec")))

# create colorblind-friendly palette
pal <- c("#56B4E9", "#D55E00", "#E69F00", "#0072B2", "#009E73", "#999999")
names(pal) <- c("amphibiens", "mammifères", "oiseaux", "poissons", "reptiles", "inconnu")

ggplot(lpd_sprich) + 
  geom_bar(aes(x = taxa, 
               y = n, 
               alpha = category,
               fill = taxa), 
           stat = "identity") +
  labs(y = "Nombre d'espèces", x = "",
       alpha = "") +
  scale_alpha_discrete(range = c(0.4, 1)) +
  scale_fill_manual(values = pal) +
  guides(fill = FALSE)
```

#### Represents only a few regions

[ same type of barplot, but predicted richness per region or biome]


### How is the Living Planet Index calculated?

[ short demo with figures]
